---
title: "NJ newspaper titles over time"
author: "Francesca Giannetti"
date: "3/4/2020"
output: html_document
---

```{r setup, include=FALSE}
if (!(require(tidyverse))) {
  install.packages("tidyverse")
}
if (!(require(tidytext))) {
  install.packages("tidytext")
}
if (!(require(plotly))) {
  install.packages("plotly")
}
if (!(require(leaflet))) {
  install.packages("leaflet")
}
if (!(require(RColorBrewer))) {
  install.packages("RColorBrewer")
}
library(tidyverse)
library(tidytext)
library(plotly)
library(leaflet)
library(RColorBrewer)

knitr::opts_chunk$set(echo = TRUE)

# read in data
nj_papers <- read.delim("nj-newspapers-geocoded.csv", sep = ",", stringsAsFactors = FALSE)
```

## Newspapers by language and decade

New Jersey newspapers grouped by language and decade of first publication.

```{r lang}
nj_papers %>% 
  group_by(language, startDecade) %>% 
  summarize(lang_total = n()) %>% 
  ggplot(aes(x=startDecade, y=lang_total, color = language)) + 
  geom_point(alpha = 0.5) +
  geom_text(aes(label = language), 
            angle = 15, 
            check_overlap = TRUE) +
  theme_minimal() +
  ggtitle("New Newspaper Titles by Language and Decade") +
  labs(x="Decades",y="Number of New Titles by Language")
```

## By Language and Decade Minus English

Let's filter out English so we can see better what is happening with the other languages. 

```{r langNoEng}
# by language and decade, minus eng
nj_papers %>% 
  filter(!language == 'eng') %>% 
  group_by(language, startDecade) %>% 
  summarize(lang_total = n()) %>% 
  ggplot(aes(x=startDecade, y=lang_total, color = language)) + 
  geom_point(alpha = 0.5) +
  geom_text(aes(label = language), 
            angle = 15, nudge_y = 0.25,
            check_overlap = TRUE) +
  theme_minimal() +
  ggtitle("New Non-English Newspaper Titles by Language and Decade") +
  labs(x="Decades",y="Number of New Titles by Language")
```

## Top words in newspaper titles

```{r}
# top words in newspaper titles, sorted by frequency
nj_papers %>% 
  unnest_tokens(word, title) %>%
  anti_join(stop_words) %>%
  group_by(start, word) %>% 
  summarize(count = n()) %>%
  arrange(desc(count)) %>%
  View()
```

## Same, sorted by start decade

```{r}
# custom stopword list
stopWords <- as_data_frame(c("the", "an", "and", "der", 
                             "die", "das", "und", "of",
                             "in","aus","dem","or", 
                             "jersey", "new", "york", "n")) %>%
  rename(word = value)

# top n words in newspaper titles, sorted by start decade
nj_papers %>% 
  unnest_tokens(word, title) %>%
  anti_join(stopWords) %>%
  group_by(startDecade, word) %>% 
  summarize(count = n()) %>%
  arrange(startDecade,desc(count)) %>%
  top_n(10) %>%
  View()
```

## Plot the top title words

```{r titlewords}
titleWords <- nj_papers %>% 
  unnest_tokens(word, title) %>%
  anti_join(stopWords) %>%
  group_by(startDecade, word) %>% 
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count)) %>%
  mutate(decadeCount = length(startDecade)) %>%
  arrange(startDecade,desc(count))

# plot top words
ggplot(titleWords) +
  aes(x=percentage,y=startDecade,label=word) + 
  geom_point(alpha=.3) + 
  geom_text(check_overlap = TRUE, nudge_x = 0.15)
```

## Percentages

Percentages of title words over time.

```{r titlewords_ts}
# by percentage
newPapers <- nj_papers %>% 
  group_by(startDecade) %>%
  summarise(newPapers = n())

titleWords <- titleWords %>%
  left_join(newPapers, by = "startDecade") %>%
  mutate(percentage = count/newPapers) %>%
  arrange(startDecade, desc(percentage))

plot <- ggplot(titleWords %>%
                 filter(startDecade >= 1800 & startDecade <= 1950) %>%
                 top_n(3) %>% 
                 filter(percentage >= .07)) +
  aes(x=startDecade, y=percentage, color = word) +
  geom_line() +
  geom_point(size = .3) +
  ggtitle("Most Used Words in New Newspaper Titles by Decade, 1800-1950") +
  labs(x="Decades",y="Percentage of Titles",fill="Word",caption="The top words used in the titles of new newspapers during the nineteenth century by decade") + 
  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=18, hjust=0.5)) +
  theme(axis.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=14)) + 
  theme(legend.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=14)) +
  theme(legend.background = element_rect(color = "#efefef")) +
  theme(plot.caption = element_text(family = "Trebuchet MS", color="#666666", size=10, hjust = 0.5, margin = margin(15, 0, 15, 0))) +
  theme(axis.text = element_text(family = "Trebuchet MS", color="#aaaaaa", face="bold", size=10)) +
  theme(panel.background = element_rect(fill = "white")) +
  theme(panel.grid.major = element_line(color = "#efefef")) +
  theme(axis.ticks = element_line(color = "#efefef"))

ggplotly(plot)
```

## By place

Only a subset of the NJ newspapers had a city of publication, but for these ones we can visualize them as a map. 

```{r njpapersmap}
decade_started <- colorBin("YlGnBu", nj_papers$startDecade, bins = 7, 
                                 na.color = "#808080")
paper_popup <- paste("<strong>Title: </strong>", 
                     nj_papers$title, "<br>",
                     "<strong>City: </strong>",
                     nj_papers$city, "<br>",
                     "<strong>Language: </strong>",
                     nj_papers$language, "<br>",
                     "<strong>Year started: </strong>",
                     nj_papers$start, "<br>",
                     "<strong>Frequency: </strong>",
                     nj_papers$frequency)

leaflet(nj_papers) %>% 
  addProviderTiles('CartoDB.Positron') %>%
  addCircleMarkers(lng = ~lon, lat = ~lat, radius = 5, stroke = FALSE,
                   fillColor = ~decade_started(startDecade),
                   popup = paper_popup)
```

